---
title: "Explore FRET Cohort LCA Classes"
author: "P R Barber"
date: "19 October 2018"
output: pdf_document
toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction

Are there any significant differences in the makeup of the FRET based LCA Classes?

#Abbreviations 

* LCA = Latent Class Analysis based on Baysian Hazards Regression.

Apart from the covariates in the original xls file from David Fisher, the following columns are available:

* FRET = The raw FRET efficiency score as a quantification of the proportion of HER3 in a dimer with HER2.
* FRETxHER3 = The FRET efficiency x pixel HER3 intensity as a measure of the total HER3 in a dimer with HER2.
* FRET.cohort = A flag to indicate this patient was in the cohort where FRET was measured from histology.
* Class.FRET.OS and PFS = Class assignment from LCA on the FRET cohort, where OS or PFS was used as the outcome.
* P1 and P2.FRET.OS and PFS = Class assignment probability for the 2 classes of Class.FRET.OS and PFS.
* Class.OS and PFS = Class assignment from LCA on the full cohort, where OS or PFS was used as the outcome.
* P1, P2 and P3.OS and PFS = Class assignment probability for the classes of Class.OS and PFS.
* RiskScore.OS and PFS = Patient risk scores calculated from the signature based on OS and PFS outcomes.
* BenefitScore.OS and PFS = Patient benefit scores calculated from the TRT cross terms of signature based on OS and PFS outcomes.

#Conclusions

* Class 2 has higher metsites.
* Class 2 has higher NEUT score.
* Class 2 has higher FRET score.
* There is a difference in OS and PFS outcomes between the classes.

There may be other differences with a lower significance, see below.

#Load data set

Load the data and separate FRET cohort.

```{r, echo=TRUE}
load(file = "COIN_Final.Rdata")
fret_cohort <- patient_data[patient_data$FRET.cohort == 1,]
```

#PFS Classes

```{r}
class1 <- fret_cohort[fret_cohort$Class.FRET.PFS == 1,]
class2 <- fret_cohort[fret_cohort$Class.FRET.PFS == 2,]

for (d in ls(pattern="class")) print(paste(d, ":  ", dim(get(d))[1], "obs. of", dim(get(d))[2], "variables"))
```

##Boxplots

Boxplots for covariates where this is possible.

```{r, fig.align='center', fig.height=2.5, fig.width=3, fig.show='hold', echo=FALSE}
par(mar=c(2.1,4.1,2.1,1.1))

significant_covars <- vector()

for(name in colnames(patient_data)){
  if(all(is.na(class1[,name]) & all(is.na(class2[,name])))) next
  tryCatch({
    boxplot(class1[,name], class2[,name], ylab = name, xaxt="n")
    axis(1, at = 1:2, labels = c("class1", "class2"), tick = FALSE)
    mtext(c(paste0("n=", sum(!is.na(class1[,name]))), paste0("n=", sum(!is.na(class2[,name])))), 
          side = 1, line = 0, at = 1:2)
    t <- t.test(class1[,name], class2[,name])
    mtext(paste0("p-value=", signif(t$p.value, 2)), side = 3, line = 0, 
          col = ifelse(t$p.value<0.05, "red", "black"))
    if(t$p.value<0.05){significant_covars <- c(significant_covars, name)}
    }, error = function(e) print(e))
}

```

##T.tests

T.tests for interesting covariates (p-value<0.05).

```{r, echo=FALSE}

for(name in significant_covars){
  tryCatch({
    t <- t.test(class1[,name], class2[,name])
    writeLines(name)
    print(t)
    }, error = function(e) e)
}

```

##Tables

Number and percentage tables for covariates where this is possible.

```{r, echo=FALSE}

for(name in colnames(patient_data)){
  if (is.factor(patient_data[,name])){
    writeLines(paste(name, "(number)"))
    print(table(patient_data$Class.FRET.PFS, patient_data[,name]))
    writeLines("")
    writeLines(paste(name, "(%)"))
    print(signif(100*table(patient_data$Class.FRET.PFS, patient_data[,name])
          /rowSums(table(patient_data$Class.FRET.PFS, patient_data[,name])), digits = 2))
    writeLines("")
    writeLines("")
  }
}

```


#OS Classes

```{r}
class1 <- fret_cohort[fret_cohort$Class.FRET.OS == 1,]
class2 <- fret_cohort[fret_cohort$Class.FRET.OS == 2,]
```

##Boxplots

Boxplots for covariates where this is possible.

```{r, fig.align='center', fig.height=2.5, fig.width=3, fig.show='hold', echo=FALSE}
par(mar=c(2.1,4.1,2.1,1.1))

significant_covars <- vector()

for(name in colnames(patient_data)){
  if(all(is.na(class1[,name]) & all(is.na(class2[,name])))) next
  tryCatch({
    boxplot(class1[,name], class2[,name], ylab = name, xaxt="n")
    axis(1, at = 1:2, labels = c("class1", "class2"), tick = FALSE)
    mtext(c(paste0("n=", sum(!is.na(class1[,name]))), paste0("n=", sum(!is.na(class2[,name])))), 
          side = 1, line = 0, at = 1:2)
    t <- t.test(class1[,name], class2[,name])
    mtext(paste0("p-value=", signif(t$p.value, 2)), side = 3, line = 0, 
          col = ifelse(t$p.value<0.05, "red", "black"))
    if(t$p.value<0.05){significant_covars <- c(significant_covars, name)}
    }, error = function(e) e)
}

```

##T.tests

T.tests for interesting covariates (p-value<0.05).

```{r, echo=FALSE}

for(name in significant_covars){
  if(all(is.na(class1[,name]) & all(is.na(class2[,name])))) next
  tryCatch({
    t <- t.test(class1[,name], class2[,name])
    writeLines(name)
    print(t)
    }, error = function(e) e)
}

```

##Tables

Number and percentage tables for covariates where this is possible.

```{r, echo=FALSE}

for(name in colnames(patient_data)){
  if (is.factor(patient_data[,name])){
    writeLines(paste(name, "(number)"))
    print(table(patient_data$Class.FRET.OS, patient_data[,name]))
    writeLines("")
    writeLines(paste(name, "(%)"))
    print(signif(100*table(patient_data$Class.FRET.OS, patient_data[,name])
          /rowSums(table(patient_data$Class.FRET.OS, patient_data[,name])), digits = 2))
    writeLines("")
    writeLines("")
  }
}

```
