---
title: "FRET Classifier and Validation"
author: "P Barber"
date: "27 September 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
NA2mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
replaceNA <- function (df, v) {
  # replace NA in each col of df with a value from the vector v
 for(i in 1:ncol(df)) { 
    c <- df[, i]
    c <- replace(c, is.na(c), v[i])
    df[, i] <- c
 }
  return(df)
}
library(filesstrings)
library(survminer)
#source("plotHRTable.R")
library(knitr)
library(kableExtra)

JNCI_pvals <- function(p){
  if(is.na(p)){
    return("")
  }
  else if(p<0.001){
    return("<0.001***")
  } 
  else if(p==0.001){
    return("0.001***")
  } 
  else if (p<0.01){
    return(paste0(format(round(p, 3), nsmall = 3), "**"))
  }
  else if (p<0.05){
    return(paste0(format(round(p, 2), nsmall = 2), "*"))
  }
  else{
    return(format(round(p, 2), nsmall = 2))
  }
}

```

#Load patient data

```{r}
load("COIN_Final.Rdata")
```

##Encode, Impute Covariates

Isolate those covariates known to be needed for Risk Signatures etc. Impute by replacing NA by the average of other patients.

```{r}
encoded_data <- with(patient_data, data.frame(SUMLES, WBC, NEUT, SHB0, FRET, FRET.cohort))

encoded_data$RAS.Mutation <- ifelse(patient_data$RAS=="Mutation", 1, 0)
encoded_data$mlivonly <- ifelse(patient_data$mlivonly=="Yes", 1, 0)
encoded_data$PIK3CA.Mutation <- ifelse(patient_data$PIK3CA=="Mutation", 1, 0)

# Get means from the 398
df <- encoded_data[encoded_data$FRET.cohort==1,]
df <- subset(df, select = -c(FRET.cohort))
trained_means <- colMeans(df, na.rm = T)

# remove FREt.cohort col
encoded_data <- subset(encoded_data, select = -c(FRET.cohort))

# Apply to all data
#encoded_data[] <- lapply(encoded_data, NA2mean)
encoded_data[] <- replaceNA(encoded_data, trained_means)
```

#Load risk signature classifier

```{r}

Class_Predictor <- 
  read.table("SP_Classifier_Score/COIN_withFRET_July16/RiskScore_formula.txt",
                        sep = '*', col.names = c("Weight", "Covariate"), 
                        stringsAsFactors = F, skip = 2, fill = T)

kable(Class_Predictor)
```

```{r}
# Load output from Saddle Point Signature and tidy up
Class_Predictor_Betas <- 
  read.table("SP_Classifier_Score/COIN_withFRET_July16/SETCV_L2/betas_with_errors_optimised.txt",
             sep = ' ', col.names = c("Covariate", "beta", "pm", "error", 
                                      "rank", "z.score", "p.value"), 
             stringsAsFactors = F)

Class_Predictor_Betas <- subset(Class_Predictor_Betas, select=-pm)
Class_Predictor_Betas$Covariate <- sub(":$", "", Class_Predictor_Betas$Covariate)
Class_Predictor_Betas$beta <- as.numeric(
  sub("^beta\\[\\d+]=", "", Class_Predictor_Betas$beta))
Class_Predictor_Betas$rank <- last_number(Class_Predictor_Betas$rank, 
                                           decimals = F, negs = F)
Class_Predictor_Betas$z.score <- last_number(Class_Predictor_Betas$z.score, 
                                           decimals = T, negs = F)
Class_Predictor_Betas$p.value <- last_number(Class_Predictor_Betas$p.value, 
                                           decimals = T, negs = F)

# Create new cols as required
Class_Predictor_Betas$description <- Class_Predictor_Betas$Covariate
Class_Predictor_Betas$description <- 
  sub("SUMLES", "Sum of longest diameter", Class_Predictor_Betas$description)
Class_Predictor_Betas$description <- 
  sub("WBC", "White blood cell count", Class_Predictor_Betas$description)
Class_Predictor_Betas$description <- 
  sub("NEUT", "Neutrophil count", Class_Predictor_Betas$description)
Class_Predictor_Betas$description <- 
  sub("SHB0", "Haemaglobin (CTC grade)", Class_Predictor_Betas$description)
#Class_Predictor_Betas$description <- 
#  sub("FRET", "HER2-HER3 FRET efficiency", Class_Predictor_Betas$description)
Class_Predictor_Betas$description <- 
  sub("mlivonly", "Liver-only metastases", Class_Predictor_Betas$description)
Class_Predictor_Betas$description <- 
  sub("PIK3CA.Mutation", "PIK3CA mutation", Class_Predictor_Betas$description)

Class_Predictor_Betas$HR <- exp(Class_Predictor_Betas$beta)
Class_Predictor_Betas$HR.U95.CI <- exp(Class_Predictor_Betas$beta+Class_Predictor_Betas$error)
Class_Predictor_Betas$HR.L95.CI <- exp(Class_Predictor_Betas$beta-Class_Predictor_Betas$error)

# Add the constant as new row
Class_Predictor_Betas <- rbind(Class_Predictor_Betas, 
                               list("Constant", NA, NA, NA, NA, 0, "Constant", 1, 1, 1))

Class_Predictor_Betas$Weight <- Class_Predictor$Weight
Class_Predictor_Betas$Weight <- str_remove(Class_Predictor_Betas$Weight, "S=\\(")
Class_Predictor_Betas$Weight <- str_remove(Class_Predictor_Betas$Weight, "\\+ \\(")
Class_Predictor_Betas$Weight <- sub("\\- \\(", "-", Class_Predictor_Betas$Weight)
Class_Predictor_Betas$Weight <- str_remove(Class_Predictor_Betas$Weight, "\\)")
Class_Predictor_Betas$Weight <- as.numeric(Class_Predictor_Betas$Weight)

mydf <- data.frame(
    Covariate = Class_Predictor_Betas$description,
    #HazardRatio = signif(Class_Predictor_Betas$HR, digits = 3),
    #HazardLower = signif(Class_Predictor_Betas$HR.L95.CI, digits = 3),
    #HazardUpper = signif(Class_Predictor_Betas$HR.U95.CI, digits = 3),
#    Pvalue = signif(Class_Predictor_Betas$p.value, digits = 2),
    Pvalue = sapply(Class_Predictor_Betas$p.value, JNCI_pvals),
    Rank = Class_Predictor_Betas$rank,
    Weight = round(Class_Predictor_Betas$Weight, digits = 5),
    stringsAsFactors=FALSE
)

mydf <- mydf[order(mydf$Rank),]

#plotHRTable(mydf, useClass = F, useWeight = T, useRank = T, xmin = -4, xmax = 5,
#            cex = 1.0, rowspacing = 0.5, col = 1,
#            hrTicks = c(0.5, 1, 2),
#            classpos = -0.8, hrpos = 1.6, pvaluepos = 3.4,
#            weightpos = 4.47, rankpos = -0.7)

# Put blank entries for the constant
mydf$Pvalue[8] = ""
mydf$Rank[8] = ""

# Reorder cols
mydf <- subset(mydf, select = c(Covariate, Rank, Pvalue, Weight))

kable(mydf, row.names=F, col.names = c("Covariate", "Rank", "p-value", "Weight"), booktabs = T) %>%
  kable_styling(latex_options = "striped")

```



#Calculate Risk Score

```{r}
patient_data$Class_Pred_Score <- calculateRiskScore(Class_Predictor, encoded_data)
```

#ROC

ROC curve for the 398 training data, these are the ones that we have ground truth FRET OS Class for. 

```{r, fig.height=4, fig.width=4}
library(pROC)

pROC_obj <- roc(patient_data$Class.FRET.OS, patient_data$Class_Pred_Score,
            smoothed = TRUE,
            # arguments for ci
            ci=FALSE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, print.thres=TRUE)

```

##Choose an optimal threshold

```{r}
optimal <- coords(pROC_obj, x="best", input="threshold", best.method="youden")
optimal
threshold = optimal[1]
```

#Classify all patients

```{r}
patient_data$Pred.Class <- ifelse(patient_data$Class_Pred_Score>threshold, 2, 1)
```

##Confusion table for the 398

```{r}

GT <- patient_data$Class.FRET.OS
Pred <- patient_data$Pred.Class

table(GT, Pred)
```

#Analyse 398 training set

```{r}
data <- patient_data[patient_data$FRET.cohort==1,]
```

```{r}
library(survival)

data$SurvObj.pfs <- with(data, Surv(pfstime, pfsevent))
data$SurvObj.os <- with(data, Surv(ostime, osevent))
col=c("red", "red", "blue", "blue") 
lty=c(1,2,1,2)

km <- survfit(SurvObj.os ~ Pred.Class + TRT, data=data)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n", cex=0.8)
print(km)

survdiff(SurvObj.os ~ Pred.Class, data=data)
survdiff(SurvObj.os[Pred.Class==1] ~ TRT[Pred.Class==1], data=data)
survdiff(SurvObj.os[Pred.Class==2] ~ TRT[Pred.Class==2], data=data)
```

```{r, fig.height=1.8, fig.width=10}
ggrisktable(km, data=data) + theme_cleantable()
```

```{r}
#Split FRET by Tertiles
breaks <- quantile(data$FRET, probs = c(0.0, 0.67, 1.0), na.rm = T)
data$FRETeff <- cut(data$FRET, breaks = breaks, labels = c("low", "high"), 
                    include.lowest = T)
data$FRET_raw <- data$FRET
data$FRET <- data$FRETeff

km <- survfit(SurvObj.os ~ Pred.Class + FRET, data=data)
plot(km, col=col, lty=lty, xlab="Overall Survival (days)", ylab="Probability of Survival")
legend("topright", col=col, legend = names(km$strata), lty=lty, bty="n", cex=0.8)
print(km)

survdiff(SurvObj.os ~ Pred.Class, data=data)

survdiff(SurvObj.os[Pred.Class==1] ~ FRET[Pred.Class==1], data=data)
survdiff(SurvObj.os[Pred.Class==2] ~ FRET[Pred.Class==2], data=data)

```

```{r, fig.height=1.8, fig.width=10}
ggrisktable(km, data=data) + theme_cleantable()
```



#Analyse COIN III 152 validation set

```{r}
data <- patient_data[patient_data$FRET.cohort==2,]
```

```{r}
library(survival)

data$SurvObj.pfs <- with(data, Surv(pfstime, pfsevent))
data$SurvObj.os <- with(data, Surv(ostime, osevent))
col=c("red", "red", "blue", "blue") 
lty=c(1,2,1,2)

km <- survfit(SurvObj.os ~ Pred.Class + TRT, data=data)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n", cex=0.8)
print(km)
```

```{r, fig.height=1.8, fig.width=10}
ggrisktable(km, data=data) + theme_cleantable()
```

```{r}
col=c("red", "blue") 
lty=c(1,2)
km <- survfit(SurvObj.os ~ Pred.Class, data=data)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n", cex=0.8)
print(km)

survdiff(SurvObj.os ~ Pred.Class, data=data)
survdiff(SurvObj.os[Pred.Class==1] ~ TRT[Pred.Class==1], data=data)
survdiff(SurvObj.os[Pred.Class==2] ~ TRT[Pred.Class==2], data=data)

```

```{r, fig.height=1.8, fig.width=10}
ggrisktable(km, data=data) + theme_cleantable()
```


##KM for raw FRET split by class

In Fig 4, FRET was split by tertiles of the original 398 patients. Use these same splits.

```{r}
col=c("red", "red", "blue", "blue") 
lty=c(1,2,1,2)

#Split FRET by Tertiles, or comment ot use same breaks as the 398
breaks <- quantile(data$FRET, probs = c(0.0, 0.67, 1.0), na.rm = T)
data$FRETeff <- cut(data$FRET, breaks = breaks, labels = c("low", "high"), 
                    include.lowest = T)
data$FRET_raw <- data$FRET
data$FRET <- data$FRETeff

km <- survfit(SurvObj.os ~ Pred.Class + FRET, data=data)
plot(km, col=col, lty=lty, xlab="Overall Survival (days)", ylab="Probability of Survival")
legend("topright", col=col, legend = names(km$strata), lty=lty, bty="n", cex=0.8)
print(km)

survdiff(SurvObj.os ~ Pred.Class, data=data)

survdiff(SurvObj.os[Pred.Class==1] ~ FRET[Pred.Class==1], data=data)
survdiff(SurvObj.os[Pred.Class==2] ~ FRET[Pred.Class==2], data=data)

```

```{r, fig.height=1.8, fig.width=10}
ggrisktable(km, data=data) + theme_cleantable()
```


#Session Information

```{r}
sessionInfo()
```
