---
title: "Explore covariates of the signature"
author: "P Barber"
date: "2 October 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
NA2mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
replaceNA <- function (df, v) {
  # replace NA in each col of df with a value from the vector v
 for(i in 1:ncol(df)) { 
    c <- df[, i]
    c <- replace(c, is.na(c), v[i])
    df[, i] <- c
 }
  return(df)
}
library(filesstrings)
library(survival)
library(survminer)
#source("plotHRTable.R")
library(knitr)
#library(kableExtra)
library(ggplot2)
```

#Load patient data

```{r}
load("COIN_Final.Rdata")
patient_data$SurvObj.pfs <- with(patient_data, Surv(pfstime, pfsevent))
patient_data$SurvObj.os <- with(patient_data, Surv(ostime, osevent))
```

##Encode, Impute Covariates

Isolate those covariates known to be needed for Risk Signatures etc. Impute by replacing NA by the average of other patients.

```{r}
encoded_data <- with(patient_data, data.frame(SUMLES, WBC, NEUT, SHB0, FRET, FRET.cohort))

encoded_data$RAS.Mutation <- ifelse(patient_data$RAS=="Mutation", 1, 0)
encoded_data$mlivonly <- ifelse(patient_data$mlivonly=="Yes", 1, 0)
encoded_data$PIK3CA.Mutation <- ifelse(patient_data$PIK3CA=="Mutation", 1, 0)

# Get means from the 398
df <- encoded_data[encoded_data$FRET.cohort==1,]
df <- subset(df, select = -c(FRET.cohort))
trained_means <- colMeans(df, na.rm = T)

# remove FREt.cohort col
encoded_data <- subset(encoded_data, select = -c(FRET.cohort))

# Apply to all data
#encoded_data[] <- lapply(encoded_data, NA2mean)
encoded_data[] <- replaceNA(encoded_data, trained_means)
```

#Load risk signature classifier

```{r}

Class_Predictor <- 
  read.table("SP_Classifier_Score/COIN_withFRET_July16/RiskScore_formula.txt",
                        sep = '*', col.names = c("Weight", "Covariate"), 
                        stringsAsFactors = F, skip = 2, fill = T)

kable(Class_Predictor)
```

```{r}
# Load output from Saddle Point Signature and tidy up
Class_Predictor_Betas <- 
  read.table("SP_Classifier_Score/COIN_withFRET_July16/SETCV_L2/betas_with_errors_optimised.txt",
             sep = ' ', col.names = c("Covariate", "beta", "pm", "error", 
                                      "rank", "z.score", "p.value"), 
             stringsAsFactors = F)

Class_Predictor_Betas <- subset(Class_Predictor_Betas, select=-pm)
Class_Predictor_Betas$Covariate <- sub(":$", "", Class_Predictor_Betas$Covariate)
Class_Predictor_Betas$beta <- as.numeric(
  sub("^beta\\[\\d+]=", "", Class_Predictor_Betas$beta))
Class_Predictor_Betas$rank <- last_number(Class_Predictor_Betas$rank, 
                                           decimals = F, negs = F)
Class_Predictor_Betas$z.score <- last_number(Class_Predictor_Betas$z.score, 
                                           decimals = T, negs = F)
Class_Predictor_Betas$p.value <- last_number(Class_Predictor_Betas$p.value, 
                                           decimals = T, negs = F)

# Create new cols as required
Class_Predictor_Betas$description <- Class_Predictor_Betas$Covariate
Class_Predictor_Betas$description <- 
  sub("SUMLES", "Sum of longest diameter", Class_Predictor_Betas$description)
Class_Predictor_Betas$description <- 
  sub("WBC", "White blood cell count", Class_Predictor_Betas$description)
Class_Predictor_Betas$description <- 
  sub("NEUT", "Neutrophil count", Class_Predictor_Betas$description)
Class_Predictor_Betas$description <- 
  sub("SHB0", "Haemaglobin (CTC grade)", Class_Predictor_Betas$description)
#Class_Predictor_Betas$description <- 
#  sub("FRET", "HER2-HER3 FRET efficiency", Class_Predictor_Betas$description)
Class_Predictor_Betas$description <- 
  sub("mlivonly", "Liver-only metastases", Class_Predictor_Betas$description)
Class_Predictor_Betas$description <- 
  sub("PIK3CA.Mutation", "PIK3CA mutation", Class_Predictor_Betas$description)

Class_Predictor_Betas$HR <- exp(Class_Predictor_Betas$beta)
Class_Predictor_Betas$HR.U95.CI <- exp(Class_Predictor_Betas$beta+Class_Predictor_Betas$error)
Class_Predictor_Betas$HR.L95.CI <- exp(Class_Predictor_Betas$beta-Class_Predictor_Betas$error)

# Add the constant as new row
Class_Predictor_Betas <- rbind(Class_Predictor_Betas, 
                               list("Constant", NA, NA, NA, NA, 0, "Constant", 1, 1, 1))

Class_Predictor_Betas$Weight <- Class_Predictor$Weight
Class_Predictor_Betas$Weight <- str_remove(Class_Predictor_Betas$Weight, "S=\\(")
Class_Predictor_Betas$Weight <- str_remove(Class_Predictor_Betas$Weight, "\\+ \\(")
Class_Predictor_Betas$Weight <- sub("\\- \\(", "-", Class_Predictor_Betas$Weight)
Class_Predictor_Betas$Weight <- str_remove(Class_Predictor_Betas$Weight, "\\)")
Class_Predictor_Betas$Weight <- as.numeric(Class_Predictor_Betas$Weight)


```


#Calculate Total Risk Score

```{r}
patient_data$Class_Pred_Score <- calculateRiskScore(Class_Predictor, encoded_data)
```


#Calculate Residual Risk Scores

```{r}
patient_data$FRET_Residual_Score <- calculateRiskScore(subset(Class_Predictor, Covariate!="FRET"), encoded_data)
patient_data$SUMLES_Residual_Score <- calculateRiskScore(subset(Class_Predictor, Covariate!="SUMLES"), encoded_data)
patient_data$WBC_Residual_Score <- calculateRiskScore(subset(Class_Predictor, Covariate!="WBC"), encoded_data)
patient_data$NEUT_Residual_Score <- calculateRiskScore(subset(Class_Predictor, Covariate!="NEUT"), encoded_data)
patient_data$SHB0_Residual_Score <- calculateRiskScore(subset(Class_Predictor, Covariate!="SHB0"), encoded_data)
patient_data$mlivonly_Residual_Score <- calculateRiskScore(subset(Class_Predictor, Covariate!="mlivonly"), encoded_data)
patient_data$PIK3CA.Mutation_Residual_Score <- calculateRiskScore(subset(Class_Predictor, Covariate!="PIK3CA.Mutation"), encoded_data)
```

#ROC

ROC curve for the 398 training data, these are the ones that we have ground truth FRET OS Class for. 

```{r, fig.height=4, fig.width=4}
library(pROC)

pROC_obj <- roc(patient_data$Class.FRET.OS, patient_data$Class_Pred_Score,
            smoothed = TRUE,
            # arguments for ci
            ci=FALSE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, print.thres=TRUE)

```

##Choose an optimal threshold

```{r}
optimal <- coords(pROC_obj, x="best", input="threshold", best.method="youden")
optimal
threshold = optimal[1]
```

#Classify all patients according to predicted LCA Class

```{r}
patient_data$Pred.Class <- ifelse(patient_data$Class_Pred_Score>threshold, 2, 1)
```

#Investigate FRET vs its Residual Score

##Get fresh data set of FRET patients

```{r}
data <- subset(patient_data, FRET.cohort!=0)
```

##Classifiy FRET patients according to raw FRET

```{r}
# Cut data and name change
data$FRETraw <- data$FRET
breaks <- quantile(data$FRETraw, probs = c(0.0, 0.5, 1.0), na.rm = T)
data$FRET <- cut(data$FRETraw, breaks = breaks, labels = c("low", "high"), include.lowest = T)
```

##Group patients according to FRET and predicted class

```{r}
data$Group <- ifelse(data$FRET == "low",
                     ifelse(data$Pred.Class == 1, "iii", "i"),
                     ifelse(data$Pred.Class == 1, "iv", "ii"))
```

##Plot FRET vs residual signature

```{r}
data <- subset(data, FRET_Residual_Score<10)   # one outlier: really high residual score

ggplot(data, aes(FRETraw, FRET_Residual_Score)) +
  geom_point(aes(color = Group, shape = TRT))
```

##Plot survival curves for the groups

```{r}
col=c("#F8766D", "#7CAe00", "#00BFC4", "#C77CFF") # ggplot default colours
lty=c(1)

km <- survfit(SurvObj.os ~ Group, data=data)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
```

```{r}
col=c("red", "blue") 
lty=c(1,2)

data_g <- subset(data, Group=="i")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="ii")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="iii")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="iv")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)


```

#Investigate SUMLES vs its Residual Score

##Get fresh data set of FRET patients

```{r}
data <- subset(patient_data, FRET.cohort!=0)
```

##Classifiy FRET patients according to SUMLES

```{r}
data$SUMLESraw <- data$SUMLES
breaks <- quantile(data$SUMLESraw, probs = c(0.0, 0.5, 1.0), na.rm = T)
data$SUMLES <- cut(data$SUMLESraw, breaks = breaks, labels = c("low", "high"), include.lowest = T)
```

##Group patients according to SUMLES and predicted class

```{r}
data$Group <- ifelse(data$SUMLES == "low",
                     ifelse(data$Pred.Class == 1, "iii", "i"),
                     ifelse(data$Pred.Class == 1, "iv", "ii"))
```

##Plot SUMLES vs residual signature

```{r}
data <- subset(data, SUMLES_Residual_Score<8)   # one outlier: really high residual score

ggplot(data, aes(SUMLESraw, SUMLES_Residual_Score)) +
  geom_point(aes(color = Group, shape = TRT))
```

##Plot survival curves for the groups

```{r}
col=c("#F8766D", "#7CAe00", "#00BFC4", "#C77CFF") # ggplot default colours
lty=c(1)

km <- survfit(SurvObj.os ~ Group, data=data)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
```

```{r}
col=c("red", "blue") 
lty=c(1,2)

data_g <- subset(data, Group=="i")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="ii")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="iii")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="iv")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)


```


#Investigate WBC vs its Residual Score

##Get fresh data set of FRET patients

```{r}
data <- subset(patient_data, FRET.cohort!=0)
```

##Classifiy FRET patients according to WBC

```{r}
data$WBCraw <- data$WBC
breaks <- quantile(data$WBCraw, probs = c(0.0, 0.5, 1.0), na.rm = T)
data$WBC <- cut(data$WBCraw, breaks = breaks, labels = c("low", "high"), include.lowest = T)
```

##Group patients according to WBC and predicted class

```{r}
data$Group <- ifelse(data$WBC == "low",
                     ifelse(data$Pred.Class == 1, "iii", "i"),
                     ifelse(data$Pred.Class == 1, "iv", "ii"))
```

##Plot WBC vs residual signature

```{r}
data <- subset(data, WBC_Residual_Score<8)   # one outlier: really high residual score

ggplot(data, aes(WBCraw, WBC_Residual_Score)) +
  geom_point(aes(color = Group, shape = TRT))
```

##Plot survival curves for the groups

```{r}
col=c("#F8766D", "#7CAe00", "#00BFC4", "#C77CFF") # ggplot default colours
lty=c(1)

km <- survfit(SurvObj.os ~ Group, data=data)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
```

```{r}
col=c("red", "blue") 
lty=c(1,2)

data_g <- subset(data, Group=="i")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="ii")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="iii")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="iv")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)


```


#Investigate NEUT vs its Residual Score

##Get fresh data set of FRET patients

```{r}
data <- subset(patient_data, FRET.cohort!=0)
```

##Classifiy FRET patients according to NEUT

```{r}
data$NEUTraw <- data$NEUT
breaks <- quantile(data$NEUTraw, probs = c(0.0, 0.5, 1.0), na.rm = T)
data$NEUT <- cut(data$NEUTraw, breaks = breaks, labels = c("low", "high"), include.lowest = T)
```

##Group patients according to NEUT and predicted class

```{r}
data$Group <- ifelse(data$NEUT == "low",
                     ifelse(data$Pred.Class == 1, "iii", "i"),
                     ifelse(data$Pred.Class == 1, "iv", "ii"))
```

##Plot NEUT vs residual signature

```{r}
#data <- subset(data, NEUT_Residual_Score<8)   # one outlier: really high residual score

ggplot(data, aes(NEUTraw, NEUT_Residual_Score)) +
  geom_point(aes(color = Group, shape = TRT))
```

##Plot survival curves for the groups

```{r}
col=c("#F8766D", "#7CAe00", "#00BFC4", "#C77CFF") # ggplot default colours
lty=c(1)

km <- survfit(SurvObj.os ~ Group, data=data)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
```

```{r}
col=c("red", "blue") 
lty=c(1,2)

data_g <- subset(data, Group=="i")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="ii")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="iii")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="iv")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)


```

#Investigate SHB0 vs its Residual Score

##Get fresh data set of FRET patients

```{r}
data <- subset(patient_data, FRET.cohort!=0)
```

##Classifiy FRET patients according to SHB0

```{r}
data$SHB0raw <- data$SHB0
data$SHB0 <- ifelse(data$SHB0raw > 0, "high", "low")
```

##Group patients according to SHB0 and predicted class

```{r}
data$Group <- ifelse(data$SHB0 == "low",
                     ifelse(data$Pred.Class == 1, "iii", "i"),
                     ifelse(data$Pred.Class == 1, "iv", "ii"))
```

##Plot SHB0 vs residual signature

```{r}
#data <- subset(data, SHB0_Residual_Score<8)   # one outlier: really high residual score

ggplot(data, aes(SHB0raw, SHB0_Residual_Score)) +
  geom_jitter(width=0.2, aes(color = Group, shape = TRT))
```

##Plot survival curves for the groups

```{r}
col=c("#F8766D", "#7CAe00", "#00BFC4", "#C77CFF") # ggplot default colours
lty=c(1)

km <- survfit(SurvObj.os ~ Group, data=data)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
```

```{r}
col=c("red", "blue") 
lty=c(1,2)

data_g <- subset(data, Group=="i")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="ii")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="iii")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="iv")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)


```

#Investigate mlivonly vs its Residual Score

##Get fresh data set of FRET patients

```{r}
data <- subset(patient_data, FRET.cohort!=0)
```

##Group patients according to mlivonly and predicted class

```{r}
data$Group <- ifelse(data$mlivonly == "No",
                     ifelse(data$Pred.Class == 1, "iii", "i"),
                     ifelse(data$Pred.Class == 1, "iv", "ii"))
```

##Plot mlivonly vs residual signature

```{r}
#data <- subset(data, mlivonly_Residual_Score<8)   # one outlier: really high residual score

ggplot(data, aes(mlivonly, mlivonly_Residual_Score)) +
  geom_jitter(width=0.2, aes(color = Group, shape = TRT))
```

##Plot survival curves for the groups

```{r}
col=c("#F8766D", "#7CAe00", "#00BFC4", "#C77CFF") # ggplot default colours
lty=c(1)

km <- survfit(SurvObj.os ~ Group, data=data)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
```

```{r}
col=c("red", "blue") 
lty=c(1,2)

data_g <- subset(data, Group=="i")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="ii")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="iii")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="iv")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)


```

#Investigate PIK3CA.Mutation vs its Residual Score

##Get fresh data set of FRET patients

```{r}
data <- subset(patient_data, FRET.cohort!=0)
```


##Group patients according to PIK3CA.Mutation and predicted class

```{r}
data$Group <- ifelse(data$PIK3CA == "Wild-type",
                     ifelse(data$Pred.Class == 1, "iii", "i"),
                     ifelse(data$Pred.Class == 1, "iv", "ii"))
```

##Plot PIK3CA.Mutation vs residual signature

```{r}
#data <- subset(data, PIK3CA.Mutation_Residual_Score<8)   # one outlier: really high residual score

ggplot(data, aes(PIK3CA, PIK3CA.Mutation_Residual_Score)) +
  geom_jitter(width=0.2, aes(color = Group, shape = TRT))
```

##Plot survival curves for the groups

```{r}
col=c("#F8766D", "#7CAe00", "#00BFC4", "#C77CFF") # ggplot default colours
lty=c(1)

km <- survfit(SurvObj.os ~ Group, data=data)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
```

```{r}
col=c("red", "blue") 
lty=c(1,2)

data_g <- subset(data, Group=="i")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="ii")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="iii")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)

data_g <- subset(data, Group=="iv")

km <- survfit(SurvObj.os ~ TRT, data=data_g)
plot(km, col=col, xlab="Overall Survival (days)", ylab="Probability of Survival", lty = lty)
legend("topright", col=col, legend = names(km$strata), lty = lty, bty="n")
print(km)
survdiff(SurvObj.os ~ TRT, data=data_g)


```


#Other plots

```{r, eval=F}
dataA <- subset(data, TRT=="A")

ggplot(dataA, aes(FRETraw, Class_Pred_Residual_Score)) +
  geom_point(aes(color = ostime, shape = as.ordered(Pred.Class))) +
  scale_colour_gradient2()

dataB <- subset(data, TRT=="B")

ggplot(dataB, aes(FRETraw, Class_Pred_Residual_Score)) +
  geom_point(aes(color = ostime, shape = as.ordered(Pred.Class))) +
  scale_colour_gradient2()


```

#Session Information

```{r}
sessionInfo()
```
